### 2020-07-04 Thesis-Chapter1-Introduction.tex
Fix usage of hyphen, en dash, and em dash

\principle{Perfection is not the standard}{Wicked policy problems do not have perfect solutions, or even verifiably
optimal solutions, so we cannot use perfection as the standard. To quote the ``insecurity axioms,'' insecurity cannot be
destroyed, only moved around \cite{nrc_schneider_1999}. \Ac{EA}, like all security problems, is a matter of finding the
right balance of risk given the threats under consideration--including the threat of regulatory action.}

@@

This thesis aims to clarify the arguments composing debate and the nature of the technical problem. After reproducing a
brief U.S.-focused history of encryption regulation, technology, and conflict, I introduce argument maps and threat
models using data flow diagrams. In the following chapters, I analyze exceptional access arguments in detail and create
a multi-level threat model, broken down by the significantly different challenges posed by \acl{DAR} and \acl{DIM}. The
thesis concludes with a set of policies and principles to apply to the problem in light of valid arguments and threats.

>>>>

\principle{Perfection is not the standard}{Wicked policy problems do not have perfect solutions, or even verifiably
optimal solutions, so we cannot use perfection as the standard. To quote the ``insecurity axioms,'' insecurity cannot be
destroyed, only moved around \cite{nrc_schneider_1999}. \Ac{EA}, like all security problems, is a matter of finding the
right balance of risk given the threats under consideration---including the threat of regulatory action.}

@@

This thesis aims to clarify the arguments composing debate and the nature of the technical problem. After reproducing a
brief U.S.--focused history of encryption regulation, technology, and conflict, I introduce argument maps and threat
models using data flow diagrams. In the following chapters, I analyze exceptional access arguments in detail and create
a multi-level threat model, broken down by the significantly different challenges posed by \acl{DAR} and \acl{DIM}. The
thesis concludes with a set of policies and principles to apply to the problem in light of valid arguments and threats.

### 2020-07-04 Thesis-Chapter2-Background.tex
Fix usage of hyphen, en dash, and em dash

% There are two classes of cryptographic ciphers--stream and block

@@

One can also define two major cryptographic applications--securing \acf{DIM} and \acf{DAR}. Each application is a
different problem that results in different solutions. \Ac{DIM} typically uses long-lived \ac{asym-crypto} keys to
perform authentication and establish short-lived \ac{sym-crypto} session keys; the session keys perform the actual
encryption of the data in transit. Two important properties of \ac{DIM} encryption protocols are \ac{forward-secrecy}
and \ac{replay-protection}. \Ac{forward-secrecy} ensures that a leaked private key or session key does not compromise
any other private key or session key; \ac{replay-protection} ensures that messages cannot be replayed by an
attacker without detection \cite{bellovin_thinking_2016}. \Ac{E2EE} for instant messaging services is an example of
encryption for \ac{DIM}.

@@

% Spoofing                  Authentication
% Tampering                 Integrity
% Repudiation               Non- Repudiation
% Information Disclosure    Confidentiality
% Denial of Service         Availability
% Elevation of Privilege    Authorization

@@

The discovery of \ac{public-key-crypto} was an important advance. Combined with computer networking and personal
computing advances in the 1980s, it for the first time put the power of strong encryption not just in the hands of
governments and militaries, but of ordinary people. However, this power enables very strong privacy, and privacy is a
property governments aim to regulate. The use of encryption for for private purposes has a contentious history. In 1587,
Mary Queen of Scots was convicted of treason based on evidence from decrypted letters, and in 1807, prosecutors trying
Aaron Burr for treason tried to force testimony from his secretary on the contents of encrypted messages
\cite{kerr_encryption_2017}. Government's natural stance towards privacy makes it inherently uncomfortable with the
rapid increase in availability of strong encryption.
% FIXME: The point is made--these developments made gov'ts uncomfortable (just look at these examples!)--but the line of
% thought is not clear enough to the reader.

@@

All these realities--government discomfort with absolute privacy, rapidly increasing availability of strong encryption,
and a blossoming computer industry foundationally reliant on cryptography--came to a head in what has become known as
the first ``crypto war.''

In 1976, the same year Diffie and Hellmann published their research in public key cryptography, the U.S. Congress passed
the \ac{AECA} and declared that strong cryptography is subject to export controls \cite{kehl_right_2015}. The opening
shots of \ac{the-cw1} came in 1991, when the U.S. Senate introduced, but did not pass, a bill mandating access to plain
text contents when authorized by law. In response, Philip Zimmermann released \ac{PGP}, a public key encryption tool to
secure email, in order that strong cryptography be ``made available to the American public before it became illegal to
use'' \cite{zimmermann_1996}. In 1993, the Clinton administration introduced the \ac{clipperchip} \cite{press_1993} with
the goal of ``providing the public with strong cryptographic tools without sacrificing the ability of law enforcement
and intelligence agencies to access unencrypted versions of those communications'' \cite{thompson_2015}. Citing the
foundational role of cryptography in security--and the potential of human rights abuses for compromised
privacy--industry and prominent technical leaders reacted negatively to the initiative \cite{kehl_right_2015}
\cite{zimmermann_1996}. When a prominent security researcher discovered flaws that allowed users to subvert the
\ac{clipperchip} mechanisms \cite{blaze_protocol_1994}, the proposal died.

@@

Two developments paved the way for \ac{the-cw2}. The first was actually a non-development: when the Clinton
administration changed its encryption policy, the U.S. House dropped the \ac{SAFEAct} \cite{goodlatte_hr3011_1996}. The
\ac{PROCODEAct} and other encryption-related bills went un-passed as well, as elaborated in \mysec{sec-us-reg-fails}.
This legislative failure meant that rather than being written in the pen of law, encryption policy was written in the
pencil of executive order.

@@

\paragraph*{``Security vs. Security''} The relationship between security and privacy is sometimes viewed as
antagonistic, an assumption implicit in in the ``nothing-to-hide'' argument, but discussions in recent years have shown
% FIXME: expand on "nothing to hide"? that the debate is more about ``Security vs. Security'' than ``Security vs.
Privacy'' \cite{stalla_bourdillon_privacy_2014}. In this context, there are two types of security: public security, the
pursuit of law enforcement, and cybersecurity, the pursuit of the technical community. Both sides of the debate have by
now acknowledged that privacy and public security are not always in conflict, and in fact \ac{EA}'s effects on public
security vs. cybersecurity--which in return affects public security--may be more important \cite{schneier_2019}.

\paragraph*{Apple vs. FBI} The 2015 San Bernardino terror attack in which the suspect was killed, leaving behind a
locked iPhone, resulted in the first major battle of \ac{the-cw2}. The \ac{FBI} issued an order under the \acrlong{AWA}
to compel Apple to unlock the device, which was among the first generation of Apple's fully encrypted iPhones; Apple
objected on grounds that it was ``unreasonably burdensome'' and would undercut the integrity of all iOS devices
\cite{schulze_clipper_2017}. The case occupied U.S. District courts, the media, and the political world's attention from
February 16, 2016, when the warrant was issued, to March 28, when the \ac{FBI} announced they had gained access to the
phone through alternate means \cite{novet_2016}. Even though the phone proved not to have any important data
\cite{schulze_clipper_2017}, the high profile case featured two characteristic elements of the new crypto
wars--terrorism and device encryption. Shortly after the case ended, a pair of Republican and Democratic senators
% FIXME: determine correct usage of -- vs ---
% (https://www.chicagomanualofstyle.org/qanda/data/faq/topics/HyphensEnDashesEmDashes/faq0002.html)
jointly released a discussion draft of a bill that would have forced Apple to comply \cite{burr_2016}. The draft did not
make it to the floor, and the unresolved nature of the court case left the legal debate unresolved.

@@

The following sections list laws which directly or indirectly affect use of encryption. Though this thesis is
U.S.-focused, it is worth briefly touching on the regulatory environment of other countries. \Ac{EA} is an international
problem, as digital technology has little respect for political borders.

@@

%     \law{1986}{\acf{CFAA}}{hughes_hr4718_1986}{
% Criminalized ``knowingly access[ing] a computer without authorization or exceeding authorized access'' and thereby
% violating the confidentiality, integrity, or availability of the system. While presumably intended to target hacking,
% the \acs{CFAA} has been used to prosecute a host of crimes, and suffers from dated language;
% % --what does it mean to ``access'' a ``computer'' loading a simple webpage may involve dozens of devices--
% whether violating the terms of service constitutes ``exceeding authorized access'' remains an open question
% \cite{wolff_computer_2016}. Importantly for \ac{EA}, the law includes carve-outs for law enforcement and intelligence
% activities.
%
% Follow up: CFAA going before supreme court: https://www.cyberscoop.com/cfaa-will-soon-day-supreme-court/
% }

@@

\Ac{CESA} was a legislative proposal by Clinton administration introduced on same day as encryption policy changes that
led to the abandonment of the \ac{SAFEAct}. Section 102 lays out the government position precisely--encryption is an
important tool for information security, but is also used to ``hide unlawful activity by terrorists, drug traffickers,
child pornographers, and other criminals'' (the \ac{horsemen}), and therefore ``[a]ppropriate means must be available to
fulfill these law enforcement objectives.'' The bill would establish voluntary-participation third-party key escrow with
``recovery agents'' that would provide law enforcement with access to \ac{plaintext}. Access would be compelled through
a variety of mechanisms, including warrantless mechanisms through \ac{FISA}. The White House failed to attract a
member of Congress to officially sponsor the bill.

@@

This thesis is U.S.-focused, but \ac{EA} technologies and policies have international impacts. The following is a very
brief account of the regulatory environment in various jurisdictions around the world based primarily on a 2016 analysis
by the Law Library of Congress \cite{acosta_government_2016}.

@@

% What are the most relevant questions? Can the government--
% - Compel access to--
%   - all data?
%   - data encrypted by a service provider? Who are covered service providers?
% - Compel decryption of--
%   - all data?
%   - data encrypted by a service provider? Who are covered service providers?
% - Perform mass collection?
% - What legal limits are in place (e.g. warrants)?
% - What oversight is in place (e.g. inspectors general, public transparency)?
% - Does the government prescribe technical approaches?
% - Is the government explicitly allowed to engage in hacking?

@@

Any scheme that relies on the fidelity and security of information held by an entity besides the data owner (in the case
of \ac{DAR}) or sender or receiver (in the case of \ac{DIM}) is a ``trusted-party'' key escrow approach. Though there is
a wide variety of implementation options within this category \cite{denning_taxonomy_1996}--secret or open protocol,
hardware-enabled or software-only, single key or split key, government or third-party escrow--the foundation of the
effectiveness of these approaches lies in the security of the trusted party. The \ac{clipperchip} is one example, being
a secret protocol hardware-enabled split key government escrow system \cite{denning_taxonomy_1996}.

@@

Research in \ac{the-cw1} included the introduction of ``translucent'' cryptography, in which an observer could surveil
some but not all communications using statistical cryptographic methods \cite{bellare_translucent_1996}. Under this
scheme, no keys are escrowed, but law enforcement could recover a predetermined fraction--say, 40\%--of plaintext data.
That fraction would be designed to balance privacy, digital security, and public security. This research, though
decidedly more experimental than key escrow proposals, was explicitly pursued to demonstrate that alternatives to key
escrow do exist \cite{bellare_translucent_1996}.

@@

Simpler methods can be used than computer hacking or legal coercion: encrypted data may exist in plaintext elsewhere, or
authorities may seize equipment while it is unlocked \cite{kerr_encryption_2017}. This approach relies on skills law
enforcement is traditionally good at--investigation and sting operations--and is used successfully. In 2017, an
international law enforcement operation used this technique to take down a large illicit online drug market
\cite{greenberg_2018}. This approach also does not address \ac{DIM}, and requires well-executed operations.

>>>>

% There are two classes of cryptographic ciphers---stream and block

@@

One can also define two major cryptographic applications---securing \acf{DIM} and \acf{DAR}. Each application is a
different problem that results in different solutions. \Ac{DIM} typically uses long-lived \ac{asym-crypto} keys to
perform authentication and establish short-lived \ac{sym-crypto} session keys; the session keys perform the actual
encryption of the data in transit. Two important properties of \ac{DIM} encryption protocols are \ac{forward-secrecy}
and \ac{replay-protection}. \Ac{forward-secrecy} ensures that a leaked private key or session key does not compromise
any other private key or session key; \ac{replay-protection} ensures that messages cannot be replayed by an
attacker without detection \cite{bellovin_thinking_2016}. \Ac{E2EE} for instant messaging services is an example of
encryption for \ac{DIM}.

@@

% Spoofing                  Authentication
% Tampering                 Integrity
% Repudiation               Non-Repudiation
% Information Disclosure    Confidentiality
% Denial of Service         Availability
% Elevation of Privilege    Authorization

@@

The discovery of \ac{public-key-crypto} was an important advance. Combined with computer networking and personal
computing advances in the 1980s, it for the first time put the power of strong encryption not just in the hands of
governments and militaries, but of ordinary people. However, this power enables very strong privacy, and privacy is a
property governments aim to regulate. The use of encryption for for private purposes has a contentious history. In 1587,
Mary Queen of Scots was convicted of treason based on evidence from decrypted letters, and in 1807, prosecutors trying
Aaron Burr for treason tried to force testimony from his secretary on the contents of encrypted messages
\cite{kerr_encryption_2017}. Government's natural stance towards privacy makes it inherently uncomfortable with the
rapid increase in availability of strong encryption.
% FIXME: The point is made---these developments made gov'ts uncomfortable (just look at these examples!)---but the line
% of thought is not clear enough to the reader.

@@

All these realities---government discomfort with absolute privacy, rapidly increasing availability of strong encryption,
and a blossoming computer industry foundationally reliant on cryptography---came to a head in what has become known as
the first ``crypto war.''

In 1976, the same year Diffie and Hellmann published their research in public key cryptography, the U.S. Congress passed
the \ac{AECA} and declared that strong cryptography is subject to export controls \cite{kehl_right_2015}. The opening
shots of \ac{the-cw1} came in 1991, when the U.S. Senate introduced, but did not pass, a bill mandating access to plain
text contents when authorized by law. In response, Philip Zimmermann released \ac{PGP}, a public key encryption tool to
secure email, in order that strong cryptography be ``made available to the American public before it became illegal to
use'' \cite{zimmermann_1996}. In 1993, the Clinton administration introduced the \ac{clipperchip} \cite{press_1993} with
the goal of ``providing the public with strong cryptographic tools without sacrificing the ability of law enforcement
and intelligence agencies to access unencrypted versions of those communications'' \cite{thompson_2015}. Citing the
foundational role of cryptography in security---and the potential of human rights abuses for compromised
privacy---industry and prominent technical leaders reacted negatively to the initiative \cite{kehl_right_2015}
\cite{zimmermann_1996}. When a prominent security researcher discovered flaws that allowed users to subvert the
\ac{clipperchip} mechanisms \cite{blaze_protocol_1994}, the proposal died.

@@

Two developments paved the way for \ac{the-cw2}. The first was actually a non-development: when the Clinton
administration changed its encryption policy, the U.S. House dropped the \ac{SAFEAct} \cite{goodlatte_hr3011_1996}. The
\ac{PROCODEAct} and other encryption-related bills went unpassed as well, as elaborated in \mysec{sec-us-reg-fails}.
This legislative failure meant that rather than being written in the pen of law, encryption policy was written in the
pencil of executive order.

@@

\paragraph*{``Security vs. Security''} The relationship between security and privacy is sometimes viewed as
antagonistic, an assumption implicit in in the ``nothing-to-hide'' argument, but discussions in recent years have shown
% FIXME: expand on "nothing to hide"? that the debate is more about ``Security vs. Security'' than ``Security vs.
Privacy'' \cite{stalla_bourdillon_privacy_2014}. In this context, there are two types of security: public security, the
pursuit of law enforcement, and cybersecurity, the pursuit of the technical community. Both sides of the debate have by
now acknowledged that privacy and public security are not always in conflict, and in fact \ac{EA}'s effects on public
security vs. cybersecurity---which in return affects public security---may be more important \cite{schneier_2019}.

\paragraph*{Apple vs. FBI} The 2015 San Bernardino terror attack in which the suspect was killed, leaving behind a
locked iPhone, resulted in the first major battle of \ac{the-cw2}. The \ac{FBI} issued an order under the \acrlong{AWA}
to compel Apple to unlock the device, which was among the first generation of Apple's fully encrypted iPhones; Apple
objected on grounds that it was ``unreasonably burdensome'' and would undercut the integrity of all iOS devices
\cite{schulze_clipper_2017}. The case occupied U.S. District courts, the media, and the political world's attention from
February 16, 2016, when the warrant was issued, to March 28, when the \ac{FBI} announced they had gained access to the
phone through alternate means \cite{novet_2016}. Even though the phone proved not to have any important data
\cite{schulze_clipper_2017}, the high profile case featured two characteristic elements of the new crypto
wars---terrorism and device encryption. Shortly after the case ended, a pair of Republican and Democratic senators
% (https://www.chicagomanualofstyle.org/qanda/data/faq/topics/HyphensEnDashesEmDashes/faq0002.html)
jointly released a discussion draft of a bill that would have forced Apple to comply \cite{burr_2016}. The draft did not
make it to the floor, and the unresolved nature of the court case left the legal debate unresolved.

@@

The following sections list laws which directly or indirectly affect use of encryption. Though this thesis is
U.S.--focused, it is worth briefly touching on the regulatory environment of other countries. \Ac{EA} is an international
problem, as digital technology has little respect for political borders.

@@

%     \law{1986}{\acf{CFAA}}{hughes_hr4718_1986}{
% Criminalized ``knowingly access[ing] a computer without authorization or exceeding authorized access'' and thereby
% violating the confidentiality, integrity, or availability of the system. While presumably intended to target hacking,
% the \acs{CFAA} has been used to prosecute a host of crimes, and suffers from dated language;
% % ---what does it mean to ``access'' a ``computer'' loading a simple webpage may involve dozens of devices---
% whether violating the terms of service constitutes ``exceeding authorized access'' remains an open question
% \cite{wolff_computer_2016}. Importantly for \ac{EA}, the law includes carve-outs for law enforcement and intelligence
% activities.
%
% Follow up: CFAA going before supreme court: https://www.cyberscoop.com/cfaa-will-soon-day-supreme-court/
% }

@@

\Ac{CESA} was a legislative proposal by Clinton administration introduced on same day as encryption policy changes that
led to the abandonment of the \ac{SAFEAct}. Section 102 lays out the government position precisely---encryption is an
important tool for information security, but is also used to ``hide unlawful activity by terrorists, drug traffickers,
child pornographers, and other criminals'' (the \ac{horsemen}), and therefore ``[a]ppropriate means must be available to
fulfill these law enforcement objectives.'' The bill would establish voluntary-participation third-party key escrow with
``recovery agents'' that would provide law enforcement with access to \ac{plaintext}. Access would be compelled through
a variety of mechanisms, including warrantless mechanisms through \ac{FISA}. The White House failed to attract a
member of Congress to officially sponsor the bill.

@@

This thesis is U.S.--focused, but \ac{EA} technologies and policies have international impacts. The following is a very
brief account of the regulatory environment in various jurisdictions around the world based primarily on a 2016 analysis
by the Law Library of Congress \cite{acosta_government_2016}.

@@

% What are the most relevant questions? Can the government---
% - Compel access to---
%   - all data?
%   - data encrypted by a service provider? Who are covered service providers?
% - Compel decryption of---
%   - all data?
%   - data encrypted by a service provider? Who are covered service providers?
% - Perform mass collection?
% - What legal limits are in place (e.g. warrants)?
% - What oversight is in place (e.g. inspectors general, public transparency)?
% - Does the government prescribe technical approaches?
% - Is the government explicitly allowed to engage in hacking?

@@

Any scheme that relies on the fidelity and security of information held by an entity besides the data owner (in the case
of \ac{DAR}) or sender or receiver (in the case of \ac{DIM}) is a ``trusted-party'' key escrow approach. Though there is
a wide variety of implementation options within this category \cite{denning_taxonomy_1996}---secret or open protocol,
hardware-enabled or software-only, single key or split key, government or third-party escrow---the foundation of the
effectiveness of these approaches lies in the security of the trusted party. The \ac{clipperchip} is one example, being
a secret protocol hardware-enabled split key government escrow system \cite{denning_taxonomy_1996}.

@@

Research in \ac{the-cw1} included the introduction of ``translucent'' cryptography, in which an observer could surveil
some but not all communications using statistical cryptographic methods \cite{bellare_translucent_1996}. Under this
scheme, no keys are escrowed, but law enforcement could recover a predetermined fraction---say, 40\%---of plaintext
data. That fraction would be designed to balance privacy, digital security, and public security. This research, though
decidedly more experimental than key escrow proposals, was explicitly pursued to demonstrate that alternatives to key
escrow do exist \cite{bellare_translucent_1996}.

@@

Simpler methods can be used than computer hacking or legal coercion: encrypted data may exist in plaintext elsewhere, or
authorities may seize equipment while it is unlocked \cite{kerr_encryption_2017}. This approach relies on skills law
enforcement is traditionally good at---investigation and sting operations---and is used successfully. In 2017, an
international law enforcement operation used this technique to take down a large illicit online drug market
\cite{greenberg_2018}. This approach also does not address \ac{DIM}, and requires well-executed operations.
